---
title: "Assignment 1"
author: "Salomé Carrasco"
date: today
format: 
  html:
    toc: true
    code-fold: true
    code-summary: "Show code"
    df-print: paged
execute:
  warning: false
  message: false
---

# Introduction

Here is a reproducible Quarto workflow where we used the linear model to predict levels of extraversion from demographic and personality variables.

```{r}
#| label: setup
#install.packages("tidyverse")
library(tidyverse)
#install.packages("psych")
library(psych)
#install.packages("broom")
library(broom)
#install.packages("gt")
library(gt)
#install.packages("performance")
library(performance)
#install.packages("see") 
#reason for addition explained in Notes
library(see)

# Load the example BFI data included with the psych package
data("bfi", package = "psych")

# Standard scoring keys supplied for the BFI data
bfi_keys <- list(
  Agreeableness = c("-A1", "A2", "A3", "A4", "A5"),
  Conscientiousness = c("C1", "C2", "C3", "-C4", "-C5"),
  Extraversion = c("-E1", "-E2", "E3", "E4", "E5"),
  Neuroticism = c("N1", "N2", "N3", "N4", "N5"),
  Openness = c("O1", "-O2", "O3", "O4", "-O5")
)
```

# Methods

## Data

We analyzed the bfi data set distributed with the `psych` package. The original data set contains `r nrow(bfi)` participants and `r ncol(bfi)` variables. The first 25 variables contain personality items measuring five broad personality dimensions; the remaining variables contain demographic information.

The five personality dimensions were Agreeableness, Conscientiousness, Extraversion, Neuroticism, and Openness. Items were scored using the scoring keys supplied with the `psych` package, including reverse-scoring where required.

```{r}

#| label: prepare-data

# Score the five BFI scales.
# scoreItems() also provides scale reliability estimates.
bfi_scoring <- psych::scoreItems(
  keys = bfi_keys,
  items = bfi,
  min = 1,
  max = 6,
  impute = "none"
)

# Convert the generated scale scores to a tibble and retain demographics.
bfi_analysis <- bfi_scoring$scores |>
  as_tibble() |>
  bind_cols(
    bfi |>
      select(gender, education, age)
  ) |>
  mutate(
    gender = factor(
      gender,
      levels = c(1, 2),
      labels = c("Male", "Female")
    )
  )

```

Scale scores represent the mean of the available items belonging to each personality dimension. Missing item responses were not imputed; scale scores were instead calculated from the available keyed items.

## Analysis

We first summarized the distributions of the five personality scores and examined their internal consistency using coefficient alpha. We then fitted two linear regression models predicting Extraversion

The first model included age and gender as predictors. The second model additionally included Neuroticism and Conscientiousness. The nested models were compared in terms of their R² and information criteria.

Before interpreting the retained model, we visually evaluated its regression assumptions using diagnostic plots produced by the `performance` package. These diagnostics were used to examine residual normality, linearity and constant residual variance, influential observations, and multicollinearity.

# Results

## Descriptive statistics and reliability

```{r}
#| label: descriptives

scale_names <- names(bfi_keys)

# psych::describe() returns a large set of descriptive statistics.
# Keep only the quantities useful for this report.
descriptive_table <- bfi_analysis |>
  select(all_of(scale_names)) |>
  psych::describe() |>
  as.data.frame() |>
  rownames_to_column("Scale") |>
  as_tibble() |>
  select(
    Scale,
    n,
    mean,
    sd,
    median,
    min,
    max,
    skew
  )

# scoreItems() provides coefficient alpha for each scored scale.
reliability_table <- tibble(
  Scale = names(bfi_keys),
  Alpha = as.numeric(bfi_scoring$alpha)
)

scale_summary <- descriptive_table |>
  left_join(reliability_table, by = "Scale")

scale_summary |>
  gt() |>
  tab_header(
    title = "Descriptive Statistics and Reliability",
    subtitle = "Big Five personality scale scores"
  ) |>
  cols_label(
    Scale = "Scale",
    n = "N",
    mean = "Mean",
    sd = "SD",
    median = "Median",
    min = "Minimum",
    max = "Maximum",
    skew = "Skewness",
    Alpha = "α"
  ) |>
  fmt_number(
    columns = c(mean, sd, median, min, max, skew, Alpha),
    decimals = 2
  ) |>
  opt_row_striping()

```

## Regression models

Before fitting the models, cases with missing values on the variables required by the larger model were removed so that both models were estimated using exactly the same observations.

```{r}

#| label: regression-data

model_data <- bfi_analysis |>
  select(
    Neuroticism,
    age,
    gender,
    Extraversion,
    Conscientiousness
  ) |>
  filter(age > 30) |> 
  drop_na()

```

The regression analyses therefore used `r nrow(model_data)` participants.

```{r}

#| label: fit-models

model_1 <- lm(
  Extraversion ~ age + gender,
  data = model_data
)

model_2 <- lm(
  Extraversion ~ age + gender + Neuroticism + Conscientiousness,
  data = model_data
)

```

```{r}

#| label: compare-models

model_comparison <- bind_rows(
  glance(model_1) |> mutate(Model = "Model 1: Age + gender"),
  glance(model_2) |> mutate(
    Model = "Model 2: Age + gender + extraversion + neuroticism"
  )
) |>
  select(Model, r.squared, adj.r.squared, AIC, BIC)

model_comparison |>
  gt() |>
  tab_header(
    title = "Comparison of Candidate Regression Models"
  ) |>
  cols_label(
    Model = "Model",
    r.squared = "R²",
    adj.r.squared = "Adjusted R²",
    AIC = "AIC",
    BIC = "BIC"
  ) |>
  fmt_number(
    columns = c(r.squared, adj.r.squared),
    decimals = 3
  ) |>
  fmt_number(
    columns = c(AIC, BIC),
    decimals = 1
  ) |>
  opt_row_striping()
```

```{r}

#| label: nested-model-test

model_test <- anova(model_1, model_2) |>
  tidy()

model_test |>
  gt() |>
  tab_header(
    title = "Nested Model Comparison"
  ) |>
  fmt_number(
    columns = where(is.numeric),
    decimals = 3
  )
```

Adding Neuroticism and Conscientiousness increased explained variance from R² = `r glance(model_1)$r.squared |> round(3)` to R² = `r glance(model_2)$r.squared |> round(3)`. The larger model also had `r ifelse(AIC(model_2) < AIC(model_1), "a lower", "a higher")` AIC than the simpler model.

For this demonstration, we therefore retain Model 2 as the candidate final model.

## Model diagnostics

Model comparison favored Model 2, so we next examined whether its assumptions appeared reasonable before interpreting its coefficients.

```{r}

#| label: model-diagnostics
#| fig-width: 9
#| fig-height: 6
#| column: page

diagnostic_plots <- performance::check_model(
  model_2,
  check = c(
    "qq",
    "linearity",
    "homogeneity",
    "outliers",
    "vif"
  ),
  panel = FALSE,
  base_size = 13
)

plot(diagnostic_plots)

```

The diagnostic plots were examined for evidence of substantial violations of the assumptions underlying the linear regression model. In particular, we looked for strong departures from residual normality, systematic patterns in the residuals, non-constant residual variance, highly influential observations, and problematic collinearity among the predictors

## Final model

```{r}

#| label: final-model-table

final_model_table <- tidy(
  model_2,
  conf.int = TRUE,
  conf.level = .95
) |>
  mutate(
    term = recode(
      term,
      `(Intercept)` = "Intercept",
      age = "Age",
      genderFemale = "Gender: Female",
      Neuroticism = "Neuroticism",
      Conscientiousness = "Conscientiousness"
    ),
    # Format p-values for presentation
    p = case_when(
      p.value < .001 ~ "< .001",
      TRUE ~ sprintf("%.3f", p.value)
    )
  )

final_model_table |>
  select(
    term,
    estimate,
    std.error,
    statistic,
    conf.low,
    conf.high,
    p
  ) |>
  gt() |>
  tab_header(
    title = "Linear Regression Predicting Extraversion",
    subtitle = paste0(
      "N = ", nobs(model_2),
      "; R² = ", round(glance(model_2)$r.squared, 3)
    )
  ) |>
  cols_label(
    term = "Predictor",
    estimate = "b",
    std.error = "SE",
    statistic = "t",
    conf.low = "95% CI Low",
    conf.high = "95% CI High",
    p = "p"
  ) |>
  fmt_number(
    columns = c(
      estimate,
      std.error,
      statistic,
      conf.low,
      conf.high
    ),
    decimals = 2
  ) |>
  opt_row_striping()

```

# Discussion

Look at those results! Incredible!

## Notes
Above is the pasted information from index.qmd

### Step 4: 
* New Quarto document created (file > new file, quarto document > title: index) After the folder tab was labeled: Untitled1. I want to save this file, but I was worried that if I name this file 'index' this would replace the provided 'index.qmd' file provided from the lab assignment. (I put that file into a folder named script nested in the assignment1 folder) I wasn't sure what to do, so before saving I copied the content from index.qmd provided for the assignment and pasted it into the source code section. When I saved the file on my computer using ctl+s I saved it as 'index'. Thankfully the duplicate file error didn't show up.
* No need for .csv file because the data comes from package psych

### Model Diagnostics error
* Rendering the first chunk of this section leads to an error:
  + Error in loadNamespace(x) : there is no package called ‘see’
* According to [Credibly Curious](https://www.njtierney.com/posts/2020-05-29-there-is-no-pkg/index.html) the solution is to install the 'see' package 

### git push error
* After confirming my account/identity, I was able to click commit with a comment. I then followed the instructions to add the 'git push...' to the Terminal. This error shows:
  + error: src refspec main does not match any
error: failed to push some refs to 'https://github.com/saca226/lab1-workflow.git'
*According to github shown in the repository at the bottom of the page, to push an existing respository from the command line in Terminal we need to copy and paste:
  +git branch -M main (the M represents 'master' that is seen on the top right corner near the refresh button. This turns to 'main' after.)
  +git push -u origin main (not just this line as it can't be first since you need to push to 'master' first.)
